---
title: "CUHK RMSC4002 Tutorial 5"
author: "Benjamin Chan"
date: "October 23, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

### Value at Risk (VaR)
#### Read in Data
```{r}
# Read in data (a CSV file) under Dataset
d <- read.csv("./../Dataset/stock_2018.csv")
d <- d[, 2:4]                          # Remove Date

class(d)                               # class: print the names of classes an object
is.matrix(d)
x <- as.matrix(d)                      # Turn it into matrix
n <- nrow(x)                           # Number of observations
```

#### Historical Simulation
Suppose that today is day $n$ and $v_i$ is the value of a market variable (stock price or index). Then the value tomorrow estimated based on the $i$-th scenario is $\hat{v}(i)=v_n\times\frac{v_i}{v_{i-1}}$ for $i=1,\dots,n$. The VaR is calculated by the quantile of the $n$ simulated values.

Suppose we spend \$40,000 on buying HSBC, \$30,000 on CLP and \$30,000 on CK on 31 Aug 2018. Mathematically speaking, let $$w=\begin{bmatrix}w_1 \\ w_2 \\ w_3 \end{bmatrix}=\begin{bmatrix} 40,000 \\ 30,000 \\ 30,000 \end{bmatrix}$$ 
be amounts of investment in stocks.

We are going to compute the 1-day VaR of this portfolio using historical simulation.
```{r}
xn <- as.vector(x[n, ])                # Select the last observation
w <- c(40000, 30000, 30000)            # Amount on each stock
p0 <- sum(w)                           # Initial total amount
ws <- w/xn                             # Number of shares bought at day n

ns <- n-1                              # Number of scenarios
hsim <- NULL                           # initialize hsim

for (i in 1:ns) {
    t <- xn*(x[i+1,]/x[i,])            # Scenario i
    hsim <- rbind(hsim,t)              # Append t to hsim
}  

is.matrix(hsim)                        # is.matrix: test if it is a matrix
dim(hsim)                              # 988x3 matrix: 988 scenarios and 3 stocks

ws                                     # Number of shares bought at day n
is.matrix(ws)
(ws <- as.matrix(ws))                  # as.matrix: turn it into a matrix

ps <- as.vector(hsim%*%ws)             # Compute simulated portfolio value

loss <- p0-ps                          # Compute loss
(VaRs <- quantile(loss, 0.99))         # 1-day 99% VaR
```

Note that the cost of the portfolio is \$`r format(sum(w), scientific = F, big.mark = ",")` on 31 Aug 2018. The 1-day 99% VaR, obtained from the 99th percentile of loss distribution, is \$`r format(round(VaRs), big.mark = ",")`.

#### Normal Model
Let $u_1$, $u_2$ and $u_3$ be daily returns of the stock HSBC, CLP and Cheung Kong respectively. Assume that $u=(u_1,u_2,u_3)'$ follows a trivariate normal distribution with mean zero (please refer to Tutorial 3 for more details). Mathematically speaking, let
$$u=\begin{bmatrix} u_1 \\ u_2 \\ u_3 \end{bmatrix}\sim N_3(0,\Sigma),$$
where $\Sigma$ is the population covariance matrix of $u$.

The change in the portfolio value is $$\Delta P=w'u=w_1u_1+w_2u_2+w_3u_3$$ with mean $$E(\Delta P)=w'E(u)=0$$ and estimated variance $$\widehat{\text{Var}}(\Delta P)=w'Sw,$$
where the sample covariance matrix $S$ is used to replace the unknown $\Sigma$ in $\text{Var}(\Delta P)=w'\Sigma w$.

The 1-day 99% VaR is given by $z_{0.99}\times\sqrt{w'Sw}$, where $z_{0.99}\approx$ `r round(qnorm(0.99), 3)` is the 99th percentile of standard normal distribution.
```{r}
# as.ts: coerce an object to a time-series
t1 <- as.ts(d$HSBC)                    # For stock HSBC (0005)
t2 <- as.ts(d$CLP)                     # For stock CLP (0002)
t3 <- as.ts(d$CK)                      # For stock Cheung Kong (0001)

# Compute daily percentage return
u1 <- (lag(t1)-t1)/t1                  # lag: compute a lagged version of a time series
u2 <- (lag(t2)-t2)/t2
u3 <- (lag(t3)-t3)/t3

u <- cbind(u1, u2, u3)                 # Combine into matrix u
head(u)

# Vectorization: faster approach than for-loop
# Each row of u is returns on a day and w is a fixed vector
# After matrix multiplication, each element of dp is
# a dot product between w and u
dp <- u%*%w                            # Delta P = w'u
dim(dp)                                # u is 988x3 and w is 3x1

(sdp <- apply(dp, 2, sd))              # Standard deviation of portfolio

# Alternatively
S <- var(u)                            # Sample covariance matrix
sqrt(w%*%S%*%w)

(VaRn <- qnorm(0.99)*sdp)              # 1-day 99% VaR
```

Note that the 1-day 99% VaR using normal model is `VaRn` (\$`r format(round(VaRn), big.mark = ",")`) which is less than `VaRs` (\$`r format(round(VaRs), big.mark = ",")`). The normality assumption is questionable since returns have a fatter tail than that of the normal distribution. Hence the `VaRn` is over-optimistic.

#### t Model
Alternatively we can model the changes in the portfolio value by Student's t-distribution. Please refer to Tutorial 2 for more details.
```{r}
ku <- sum((dp/sdp)^4)/nrow(dp) - 3     # Sample excess kurtosis
(v <- round(6/ku + 4))                 # Estimate df, round to the nearest integer

(VaRt <- qt(0.99, v)*sdp)              # 1-day 99% VaR
```

Note that the 1-day 99% VaR using t model is `VaRt` (\$`r format(round(VaRt), big.mark = ",")`) which is greater than `VaRn` (\$`r format(round(VaRn), big.mark = ",")`).

### Backtesting
Let 1-day $X$% VaR be \$$V$. An exception occurs if the loss is greater than \$$V$ on a given day. If the VaR model is accurate, the probability that the loss is greater than \$$V$ on any given day is $p=1-X$. Suppose the total number of days is $n$. Using Binomial($n,p$), the expected number of exceptions is given by $np=n(1-X)$.
```{r}
n <- nrow(d)-1                         # Number of observation of u
n1 <- n-250+1                          # Starting index for 250 days before n
x <- as.matrix(d[n1:n,])               # Select the most recent 250 days
ps <- as.vector(x%*%ws)                # Compute portfolio values
ps <- c(ps, sum(w))                    # Add total amount at the end
loss <- ps[1:250] - ps[2:251]          # Compute the daily losses

(expected_exc <- 250*(1-0.99))         # Expected number of exceptions

# Count the number of exceptions in each VaR model
sum(loss>VaRs)                         # Historical Simulation
sum(loss>VaRn)                         # Normal Model 
sum(loss>VaRt)                         # t Model
```

From the output, the number of exceptions in the past 250 days for all models are small.