---
title: "CUHK RMSC4002 Tutorial 3"
author: "Benjamin Chan"
date: "October 9, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

### Read in and Manipulate Data
The file `stock_2018.csv` contains information about the stock HSBC (0005), CLP (0002) and Cheung Kong (0001) from 1 Sep 2014 to 31 Aug 2018. Data are downloaded at Yahoo Finance (https://finance.yahoo.com/).
```{r}
# Read in data (a CSV file) under Dataset
d <- read.csv("./../Dataset/stock_2018.csv")

# as.ts: coerce an object to a time-series
t1 <- as.ts(d$HSBC)        # For stock HSBC (0005)
t2 <- as.ts(d$CLP)         # For stock CLP (0002)
t3 <- as.ts(d$CK)          # For stock Cheung Kong (0001)

# Compute daily percentage return
u1 <- (lag(t1)-t1)/t1      # lag: compute a lagged version of a time series
u2 <- (lag(t2)-t2)/t2
u3 <- (lag(t3)-t3)/t3

u <- cbind(u1, u2, u3)     # Combine into matrix u
```

### Check for Multivariate Normal Distribution
* For univariate normal distribution $N(\mu,\sigma^2)$, the term 
$$\bigg(\frac{x-\mu}{\sigma}\bigg)^2=(x-\mu)(\sigma^2)^{-1}(x-\mu)$$
measures the square of the distance from $x$ to $\mu$ in standard deviaton units.

* For multivariate normal distribution $N_p(\mu,\Sigma)$, the term
$$(x-\mu)'\Sigma^{-1}(x-\mu)$$
is the square of the generalized distance from $x$ to $\mu$, or the Mahalanobis distance.

* If $u=(u_1,u_2,u_3)'$ follows a multivariate normal distribution, then the quadratic form $$d^2=(u-\bar{u})'S^{-1}(u-\bar{u}),$$
where $\bar{u}$ is the sample mean vector and $S$ is the sample covariance matrix, follows approximately $\chi^2_3$.

* In general, if $u=(u_1,\dots,u_p)'\sim N_p(\mu_u, \Sigma_u)$, then $d^2$ should follow approximately a chi-squared distribution with $p$ degrees of freedom.
```{r}
n2 <- nrow(u)                # Number of rows in u
n1 <- n2-180+1               # Starting index: 180th obs before n2
u180 <- u[n1:n2,]            # Save the most recent 180 days to u180

(m <- apply(u180, 2, mean))  # Compute mean vector of u180
# Transform m into 180x3 matrix (each row is the column mean)
m <- matrix(m, nr = 180, nc = 3, byrow = T)	

(s <- var(u180))             # Compute covariance matrix of u180
sinv <- solve(s)             # Compute inverse of s

dim(u180)                    # 180x3 matrix
dim(m)                       # 180x3 matrix

# Compute the squared generalized distance
d2 <- diag((u180-m) %*% sinv %*% t(u180-m))
length(d2)                   # Length of d2 is 180
```

Remark: u180 and m are 180x3 matrices while sinv is a 3x3 matrix. Hence (u180-m) %\*% sinv %\*% t(u180-m) is a 180x180 matrix. Its diagonal is a vector of length 180 equal to the squared generalized distance from each observation to the sample mean.

```{r, fig.height = 5, fig.width = 5, fig.align = "center"}
d2 <- sort(d2)               # Sort d2 in ascending order
i <- ((1:180)-0.5)/180       # Create a vector of percentiles
q <- qchisq(i, 3)            # Compute quantile of chisq(3)

par(mfrow = c(1, 1))
qqplot(q, d2)                # QQ-chisquare plot
abline(lsfit(q, d2))         # Add least squares fit line

ks.test(d2, pchisq, 3)       # KS chisquare test for d2
```

From the plot, the distribution of u180 is close to multivariate normal. However, the p-value is small, so formally speaking, you should reject the null hypothesis that d2 comes from a chi-squared distribution.

Famous statistician George Box said "All models are wrong but some are useful". Let us assume that $u=(u_1,u_2,u_3)'$ follows a trivariate normal distribution for the time being.

### Generate Multivariate Normal Random Vector
* Note that any linear combination of normal random variables/vectors are also normally distributed.

* For univariate normal distribution, if $$Z\sim N(0,1),$$ then $$X=\mu+\sigma Z\sim N(\mu,\sigma^2).$$
Reasoning: $$E(X)=\mu+\sigma E(Z)=\mu$$ and $$\text{Var}(X)=\sigma^2 \text{Var}(Z)=\sigma^2.$$

* For multivariate normal distribution, if $$Z=(Z_1,\dots,Z_p)'\sim N_p(0_p, I_p), \hspace{2mm}\text{i.e.}\hspace{2mm} Z_i\overset{\text{i.i.d.}}{\sim} N(0,1)$$ and $$C'C=\Sigma,$$ then $$X=\mu+C'Z\sim N_p(\mu,\Sigma).$$
Reasoning: $$E(X)=\mu+C'E(Z)=\mu$$ and $$\text{Cov}(X)=C'\text{Cov}(Z)C=C'C=\Sigma.$$

* The matrix $C$ is called the Cholesky decomposition of $\Sigma$. See https://en.wikipedia.org/wiki/Cholesky_decomposition for more details.
```{r}
# Setting random seed in simulation ensures that 
# the same set of pseudo random numbers is generated each time.
# It is an important element in reproducible research.
set.seed(7)                  # Set random seed
mu <- apply(u180, 2, mean)   # Compute mean vector
sigma <- var(u180)           # Compute covariance matrix
C <- chol(sigma)             # Cholesky decomposition of sigma

# Sanity check: C'C = sigma
t(C)%*%C
sigma

Tm <- cbind(t1,t2,t3)        # Combine t1, t2, t3 to form Tm
T0 <- Tm[nrow(Tm),]          # Set T0 to be the most recent prices

# Simulate prices for the future 90 days
for (i in 1:90) {             
    Z <- rnorm(3)            # Generate normal random vector
    v <- mu + t(C)%*%Z       # Transform to multivariate normal
    T1 <- T0*(1 + v)         # Predict new stock prices
    Tm <- rbind(Tm, t(T1))   # Append T1 to Tm
    T0 <- T1                 # Update T0
}
```

### Plot Simulation Results
```{r, warning = FALSE, message = FALSE}
library(plotly)              # Create Interactive Web Graphics via 'plotly.js'
library(tidyr)               # Easily Tidy Data with 'spread()' and 'gather()' Functions
library(dplyr)               # A Grammar of Data Manipulation

colnames(Tm) <- c("HSBC", "CLP", "CK")     # To be appeared in plot

# %>%: pipe operator
newseries <- as.data.frame(Tm) %>% 
    gather(type, value) %>%
    mutate(time = rep(time(Tm), 3))

# Alternatively
temp1 <- as.data.frame(Tm)
temp2 <- gather(temp1, type, value)
temp3 <- mutate(temp2, time = rep(time(Tm), 3))

head(newseries)
head(temp3)

tail(newseries)
tail(temp3)

plot_ly(x = newseries$time, y = newseries$value, color = newseries$type, mode = 'lines')
```