---
title: "CUHK RMSC4002 Tutorial 2"
author: "Benjamin Chan"
date: "October 2, 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

### Read in Data
The files `0005.HK.csv`, `0002.HK.csv` and `0001.HK.csv` contain information about the stock HSBC (0005), CLP (0002) and Cheung Kong (0001) from 1 Sep 2014 to 31 Aug 2018 respectively. Data are downloaded at Yahoo Finance (https://finance.yahoo.com/).
```{r}
# Read in data (CSV files) under Dataset
# Save them to objects named d5, d2, d1
d5 <- read.csv("./../Dataset/0005.HK.csv")
d2 <- read.csv("./../Dataset/0002.HK.csv")
d1 <- read.csv("./../Dataset/0001.HK.csv")

# as.character: coerce its argument to character type
d5$Date <- as.character(d5$Date)
d2$Date <- as.character(d2$Date)
d1$Date <- as.character(d1$Date)

str(d5)            # str: display the structure of an object
str(d2)
str(d1)
```

### Clean Data
Combine the three CSV files into one containing the adjusted daily closing prices only.
```{r}
HSBC <- d5$Adj.Close
CLP <- d2$Adj.Close
CK <- d1$Adj.Close

# Sanity check: the same date
mean(d5$Date == d2$Date)          # 1 means they are identical
mean(d2$Date == d1$Date)          # 1 means they are identical

Date <- d5$Date                           # Save the date

d <- data.frame(Date, HSBC, CLP, CK)      # data.frame: create a data frame
head(d)                                   # Return the first part of data
tail(d)                                   # Return the last part of data

# write.csv: create a CSV file
write.csv(d, "./../Dataset/stock_2018.csv", row.names = FALSE) 
```

### Time Series Plot
```{r}
is.ts(d$HSBC)              # is.ts: test whether an object is a time series
class(d$HSBC)              # class: prints the names of classes an object inherits from

# as.ts: coerce an object to a time-series
t1 <- as.ts(d$HSBC)        # For stock HSBC (0005)
is.ts(t1)
class(t1)

t2 <- as.ts(d$CLP)         # For stock CLP (0002)
t3 <- as.ts(d$CK)          # For stock Cheung Kong (0001)

# Compute daily percentage return
u1 <- (lag(t1)-t1)/t1      # lag: compute a lagged version of a time series
u2 <- (lag(t2)-t2)/t2
u3 <- (lag(t3)-t3)/t3

# Sanity check: match positive and negative returns
head(d)                    # Return the first part of data (default: 6 rows)

# First element corresponds to daily return on the 2nd day
head(u1, 5)                # Return the first 5 daily returns of HSBC
head(u2, 5)                # Return the first 5 daily returns of CLP
head(u3, 5)                # Return the first 5 daily return of Cheung Kong

# par: set or query graphical parameters
# mfrow and mfcol: the form c(nr, nc) means figures are drawn in
# an nr-by-nc array on the device by columns (mfcol) or rows (mfrow)
par()$mfrow                # mfrow = c(nr, nc)
par(mfrow = c(3,1))        # Set multi-frame for ploting
plot(u1)
plot(u2)
plot(u3)

# Compute average daily percentage return
cat("Average daily percentage return of", "\n",
    "HSBC: ", paste0(round(mean(u1)*100, 3), "% ;"), "\n",
    "CLP:  ", paste0(round(mean(u2)*100, 3), "% ;"), "\n",
    "CK:   ", paste0(round(mean(u3)*100, 3), "%"),
    sep = "")
```

As you can see, the daily percentage returns fluctuate around 0 (more precisely on upward trend).

### Histogram and Normal QQ plot
```{r}
par(mfrow = c(3,2))
hist(u1)                   # Produce histogram
qqnorm(u1)                 # Produce normal QQ plot
qqline(u1)                 # Add a line through theoretical 1st and 3rd quartiles

hist(u2)
qqnorm(u2)
qqline(u2)

hist(u3)
qqnorm(u3)
qqline(u3)
```

There is some discrepancy, especially in the two tails. It suggests that normal distribution may not be plausible. More specifically, heavier-tailed distributions are needed.

### Check for Normal Distribution
In statistical hypothesis testing, the p-value is the probability for a given statistical model that, when the null hypothesis is true, the test statistic would be the same as or of greater magnitude than the actual observed results. You would reject the null hypothesis when the p-value is small.
```{r}
# ks.test: perform a Kolmogorov-Smirnov test
ks.test(u1, pnorm)                # KS normality test for u1, u2 and u3
ks.test(u2, pnorm)
ks.test(u3, pnorm)

# Write a function for JB-test
JB.test <- function(u) {
    n <- length(u)                # Sample size
    s <- sd(u)                    # Standard deviation
    sk <- sum(u^3)/(n*s^3)        # Skewness
    ku <- sum(u^4)/(n*s^4) - 3    # Excess kurtosis
    JB <- n*(sk^2/6 + ku^2/24)    # JB test statistic
    p <- 1 - pchisq(JB, 2)        # p-value
    cat("JB-stat:", JB, " p-value:", p,"\n")
}

JB.test(u1)
JB.test(u2)
JB.test(u3)
```

Since the p-values are very small (close to 0), you should reject the null hypothesis that the daily percentage returns come from normal distribution. It matches with the intuition gained from the plots.

### Check for t-distribution
```{r}
# Write a function for QQ-t plot
QQt.plot <- function(u) {
    su <- sort(u)                     # Sort u
    n <- length(u)                    # Sample size
    s <- sd(u)                        # Standard deviation
    ku <- sum(u^4)/(n*s^4) - 3	      # Excess kurtosis
    v <- round(6/ku + 4)              # Estimate df, round to the nearest integer
    i <- ((1:n) - 0.5)/n              # Create a vector of percentile
    q <- qt(i, v)                     # Percentile point from t(v)
  
    hist(u)                           # Histogram of u
    plot(q, su, main = "QQ-t plot")   # Plot(q, su)
    abline(lsfit(q, su))              # Add least squares fit line
    v                                 # Output estimated df
}

par(mfrow = c(3,2))
v1 <- QQt.plot(u1)                    # QQ-t plot for u1, u2, u3 and save df
v2 <- QQt.plot(u2)
v3 <- QQt.plot(u3)

ks.test(u1, pt, v1)                   # KS t-test for u1, u2 and u3
ks.test(u2, pt, v2)
ks.test(u3, pt, v3)
```

From the plots, t-distribution fits the tails better than normal distribution does. However, the p-values are very small (close to 0) so you should reject the null hypothesis that the daily percentage returns come from t-distribution.

Famous statistician George Box said "All models are wrong but some are useful".